---
title: ''
mainfont: Arial
fontsize: 12pt
fig_width: 9
fig_height: 3.5
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 6}
   \fancyfoot[C]{\rmfamily\color{headergrey}Model Selection}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(car, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(class, quietly = TRUE, warn.conflicts = FALSE)
library(boot, quietly = TRUE, warn.conflicts = FALSE)
library(leaps, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(pls, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

select <- dplyr::select
```

## Chapter 6

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

### Lab

```{r}
hitters <- as.data.table(ISLR::Hitters)

# ggpairs(hitters)
```

```{r}
sum(is.na(hitters$Salary))
full <- nrow(hitters)
```

```{r}
hitters <- hitters[ complete.cases(hitters), ]

1 - (nrow(hitters) / full) # drop about 18% of the data

regfit.full <- regsubsets(Salary ~., hitters)
summary(regfit.full)
```

```{r}
regfit.full <- regsubsets(Salary ~., data = hitters, nvmax = 19)
reg.summary <- summary(regfit.full)

names(reg.summary)
```

```{r}
reg.summary$rsq
```

```{r}
par(mfrow = c(2,2))
plot(reg.summary$rss, xlab = "Number of Variables", ylab = "RSS")
which.min(reg.summary$rss)
points(19, reg.summary$rss[19], col="red", cex=2, pch=20)
plot(reg.summary$adjr2, xlab = "Number of Variables", ylab = "Adj. R^2")
which.max(reg.summary$adjr2)
points(11, reg.summary$adjr2[11], col="red", cex=2, pch=20)
plot(reg.summary$cp, xlab = "Number of Variables", ylab = "Cp", type = "l")
which.min(reg.summary$cp)
points(10, reg.summary$cp[10], col="red", cex=2, pch=20)
which.min(reg.summary$bic)
plot(reg.summary$bic, xlab = "Number of Variables", ylab = "BIC", type = "l")
points(6, reg.summary$bic[6], col = "red", cex = 2, pch = 20)
```

```{r}
plot(regfit.full, scale = "r2")
plot(regfit.full, scale = "adjr2")
plot(regfit.full, scale = "Cp")
plot(regfit.full, scale = "bic")
```

```{r}
coef(regfit.full, 6)
```

```{r}
regfit.fwd <- regsubsets(Salary ~., data = hitters, nvmax = 19)
summary(regfit.full)
```

```{r}
regfit.bwd <- regsubsets(Salary ~., data = hitters, nvmax = 19)
summary(regfit.bwd)
```

Cross-Validation using best subset selection:

```{r}
train <- sample(c(T, F), size = nrow(hitters), replace = T)
test <- (!train)
```

```{r}
regfit.best <- regsubsets(Salary ~., data = hitters[train,], nvmax = 19)
test_mat <- model.matrix(Salary ~., data = hitters)

val_errors <- rep(NA, 19)
for(i in 1:19)
{
   coefi <- coef(regfit.best, id = i)
   pred <- test_mat[, names(coefi)] %*% coefi
   val_errors[i] <- mean((hitters$Salary[test] - pred)^2)
}

which.min(val_errors)
```

```{r}
predict_regsubsets <- function(object, newdata, id, ...)
{
   form <- as.formula(object$call[[2]])
   mat <- model.matrix(form, newdata)
   coefi <- coef(object, id = id)
   xvars <- names(coefi)
   mat[, xvars] %*% coefi
}

regfit.best <- regsubsets(Salary ~ ., data = hitters, nvmax = 19)
coef(regfit.best, 10)
```

```{r}
k <- 10
set.seed(1)
folds <- sample(1:k, nrow(hitters), replace = T)
cv.errors <- matrix(NA, k, 19, dimnames = list(NULL, paste(1:19)))

for(j in 1:k)
{
   best.fit <- regsubsets(Salary ~ ., data = hitters[folds != j, ], nvmax = 19)
   for(i in 1:19) {
      pred <- predict_regsubsets(best.fit, hitters[folds == j, ], id = i)
      cv.errors[j, i] <- mean( (hitters$Salary[folds == j] - pred)^2)
   }
}

mean.cv.errors <- apply(cv.errors, 2, mean)
mean.cv.errors

par(mfrow = c(1,1))
plot(mean.cv.errors, type ='b')
```

### Ridge Regression

```{r}
x <- model.matrix(Salary ~ ., hitters)[, -1]
y <- hitters$Salary

grid <- 10^seq(10, -2, length = 100)

ridge.mod <- glmnet(x, y, alpha = 0, lambda = grid)
plot(ridge.mod)

dim(coef(ridge.mod))
```

```{r}
ridge.mod$lambda[50]
coef(ridge.mod)[,50]

sqrt(sum(coef(ridge.mod)[-1, 50])^2)
```

```{r}
index <- 60
ridge.mod$lambda[index]
coef(ridge.mod)[, index]

sqrt(sum(coef(ridge.mod)[-1, index])^2)
```

```{r}
set.seed(1)
train <- sample(1:nrow(x), nrow(x)/2)
test <- (-train)

y.test <- y[test]

ridge.mod <- glmnet(x[train,], y[train], alpha = 0, lambda = grid, threshold = 1e-12)

ridge.pred <- predict(ridge.mod, s = 4, newx = x[test,])
mean((ridge.pred - y.test)^2)

```

### Lasso

```{r}
lasso.mod <- glmnet(x[train,], y[train], alpha = 1, lambda = grid)
plot(lasso.mod)
```

```{r}
set.seed(1)

cv.out <- cv.glmnet(x[train, ], y[train], alpha = 1)

plot(cv.out)

bestlam <- cv.out$lambda.min
lasso.pred <- predict(lasso.mod, s = bestlam, newx = x[test,])

mean((lasso.pred - y.test)^2)
```

```{r}
out <- glmnet(x, y, alpha = 1, lambda = grid)
lasso.coef <- predict(out, type = "coefficients", s = bestlam)[1:20,]
lasso.coef
```

### Principal Components Regression

```{r}
set.seed(2)

pcr.fit <- pcr(Salary ~., data = hitters, scale = T, validation = "CV")

summary(pcr.fit)

validationplot(pcr.fit, val.type = "MSEP")
```

```{r}
set.seed(1)

pcr.fit <- pcr(Salary ~ ., data = hitters, subset = train, scale = T,
               validation = "CV")
validationplot(pcr.fit, val.type = "MSEP")
```

```{r}
pcr.pred <- predict(pcr.fit, x[test,], ncomp = 7)
mean((pcr.pred - y.test)^2)
```

### Partial Least Squares

```{r}
set.seed(1)

pls.fit <- plsr(Salary ~ ., data = hitters, subset = train, scale = T,
                validation = "CV")

summary(pls.fit)

pls.pred <- predict(pls.fit, x[test, ], ncomp = 2)
mean((pls.pred - y.test)^2)

pls.fit <- plsr(Salary ~ ., data = hitters, scale = T, ncomp = 2)
summary(pls.fit)
```


## Conceptual

### 1.)

We perform best subset, forward stepwise, and backward stepwise selection on a single data set. For each approach, we obtain $p + 1$ models, containing $0, 1, 2, \ldots, p$ predictors.

a.) Which of the three models with k predictors has the smallest *training* RSS?

_The model with the smallest training RSS will be the $C{k \choose p}$ model with p = k._

b.) Which of the three models with k predictors has the smallest *test* RSS?

_Difficult to say. The best subset selection technique looks at more models, however, forward or backward selection could pick a better model by chance._

c.) T/F

i.) The predictors in the k-variable model identified by foward stepwise selection are a subset of the predictors in the (k+1)-variable model identified by forward stepwise selection.
 
__T__

ii.) The predictors isn the k-variable model identified by backward stepwise are a subset of the predictors in the (k+1)-variable model identified by backward stepwise selection.

__T__

iii.) The predictors in the k-variable model identified by backward stepwise are a subset of the predictors in the (k+1)-variable model identified by forward stepwise selection.

__F__. There is no link between these models predictors.

iv.) The predictors in the k-variable model identifed by forward stepwise are a subset of the predictors in the (k+1)-variable model identifed by backward stepwise selection.

__F__. There is no link between these models predictors.

v.) The predictors in the k-variable model identified by best subset are a subset of the predictors in the (k+1)-variable model identifed by best subset selection.

__F__. There is no link between these models predictors.

### 2.)

For parts (a) through (c), indicate which of i through iv. is correct.

a.) The lasso, relative to least squares, is:

i.) More flexible and hence will give improved prediction accuracy when its increase in bias is less than its decrease in variance.

__T__

ii.) More flexible and hence will give improved prediction accuracy when its increase in variance is less than its decrease in bias.

__F__

iii.) Less flexible and hence will give improved prediction accuracyt when its increase in bias is less than its decrease in variance.

__F__

iv.) Less flexible and hence will give improved prediction accuracy when its increase in variance is less than its decrease in bias.

__F__

### 3.)

Suppose we estimate the regression coefficients in a linear regression model by minimizing:

$\sum^n_{i=1}{(y_i - \beta_0 - \sum^p_{j=1}\beta_jx_{i,j})}$

subject to,

$\sum^p_{j=1}(|\beta_j|) \le s$

a.) As we increase s from 0, the training RSS will:

Steadily decrease. As we increase s from 0, we are restricting the βj coefficients less and less (the coefficients will increase to their least squares estimates), and so the model is becoming more and more flexible which provokes a steady decrease in the training RSS.

b.) Test RSS will:

Decrease initially, and then eventually start increasing in a U shape. As we increase s from 0, we are restricting the βj coefficients less and less (the coefficients will increase to their least squares estimates), and so the model is becoming more and more flexible which provokes at first a decrease in the test RSS before increasing again after that in a typical U shape.

c.) Variance will:

Steadily increase. As we increase s from 0, we are restricting the βj coefficients less and less (the coefficients will increase to their least squares estimates), and so the model is becoming more and more flexible which provokes a steady increase in variance.

d.) (squared) bias will:

Steadily decrease. As we increase s from 0, we are restricting the βj coefficients less and less (the coefficients will increase to their least squares estimates), and so the model is becoming more and more flexible which provokes a steady decrease in bias.

e.) irreducible error will:

remain unchanged.

### 4.)

Suppose we estimate the regression coefficients in a linear regression model by minimizing:

$\sum^n_{i=1}(y_i - \beta_0 - \sum^p_{j=1}\beta_jx_{ij}) + \lambda\sum^p_{j=1}\beta^2_j$

for a particular value of $\lambda$.

a.) As we increase $\lambda$ from 0, the training RSS will:

Steadily increase.

b.) Test RSS will:

Decrease initially, then turn to a U shape.

c.) Variance will:

Steadly decrease.

d.) (squared) bias will:

Steadily increase

e.) Irreducible error will:

Remain constant.

### 5.)

It is well-known that ridge regression tends to give similar coefficient values to correlated variables, whereas the lasso may give quite different coefficient values to correlated variables. We will now explore this property in a very simple setting.

Suppose that $n = 2, p = 2, x_{11} = x_{12} = x_{21} = x_{22}$. Furthermore, suppose that $y_1 + y_2 = 0$ and $x_{11} + x_{21} = 0$ so the estimate for the intercept in a least squares, ridge regresion or lasso model is zero: $\hat{\beta}_0 = 0$

Write out the ridge regression optimization problem in this setting:

$(y_1 - \hat{\beta_1}x_1 - \hat{\beta_2})^2 + (y_2 - \hat{\beta_1}x_2 - \hat{\beta_2}x_2)^2 + \lambda(\hat{\beta^2_1} + \hat{\beta^2_2})$

## Applied

### 8.)

#### 9.)

#### 10.)

#### 11.)
