---
title: ''
mainfont: Arial
fontsize: 12pt
fig_width: 9
fig_height: 3.5
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 7}
   \fancyfoot[C]{\rmfamily\color{headergrey}Beyond Linearity}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(car, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(class, quietly = TRUE, warn.conflicts = FALSE)
library(boot, quietly = TRUE, warn.conflicts = FALSE)
library(leaps, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(pls, quietly = TRUE, warn.conflicts = FALSE)
library(splines, quietly = TRUE, warn.conflicts = FALSE)
library(gam, quietly = TRUE, warn.conflicts = FALSE)
library(akima, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

select <- dplyr::select

```

## Chapter 7

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Lab

### Polynomial Functions and Cut Points

```{r}
load(paste0(here::here(), "/ISLR/7.R.RData"))

plot(x, y)

fit <- lm(y ~ x)
fit2 <- lm(y ~ 1 + x + I(x^2))
```

```{r}
wage <- data.table(ISLR::Wage)
```

### Polynomial Regression and Step Functions

```{r}
fit <- lm(wage ~ poly(age, 4), data = wage)

summary(fit)

plot(fit)

coef(summary(fit))
```

```{r}
fit2 <- lm(wage ~ poly(age, 4, raw = T), data = wage)
coef(summary(fit2))
```

Alternative:

```{r}
fit2a <- lm(wage ~ age + I(age^2) + I(age^3) + I(age^4), data = wage)
coef(summary(fit2a))
```

```{r}
fit2b <- lm(wage ~ cbind(age, age^2, age^3, age^4), data = wage)
coef(fit2b)
```

```{r}
agelims <- range(wage$age)
age.grid <- seq(from = agelims[1], to = agelims[2])

pred <- predict(fit, newdata = list(age = age.grid), se = T)

se.bands <- cbind(pred$fit + 2*pred$se.fit, pred$fit - 2*pred$se.fit)

par(mfrow = c(1, 1), mar = c(4.5, 4.5, 1, 1), oma = c(0, 0, 4, 0))
plot(wage$age, wage$wage, xlim = agelims, cex = .5, col = "darkgrey")
title("Degree-4 Polynomial", outer = T)
lines(age.grid, pred$fit, lwd = 2, col = "blue")
matlines(age.grid, se.bands, lwd = 1, col = "blue", lty = 3)
```

```{r}
pred2 <- predict(fit2, newdata = list(age = age.grid), se = T)
max(abs(pred$fit - pred2$fit))
```

```{r}
fit1 <- lm(wage ~ age, data = wage)
fit2 <- lm(wage ~ poly(age, 2), data = wage)
fit3 <- lm(wage ~ poly(age, 3), data = wage)
fit4 <- lm(wage ~ poly(age, 4), data = wage)
fit5 <- lm(wage ~ poly(age, 5), data = wage)

anova(fit1, fit2, fit3, fit4, fit5)

coef(summary(fit5))
```

```{r}
fit1 <- lm(wage ~ education + age, data = wage)
fit2 <- lm(wage ~ education + poly(age, 2), data = wage)
fit3 <- lm(wage ~ education + poly(age, 3), data = wage)

anova(fit1, fit2, fit3)
```

```{r}
fit <- glm(I(wage > 250) ~ poly(age, 4), data = wage, family = "binomial")

pred <- predict(fit, newdata = list(age = age.grid), se = T)

pfit <- exp(pred$fit) / (1 + exp(pred$fit))

se.bands.logit <- cbind(pred$fit + 2 * pred$se.fit, pred$fit - 2*pred$se.fit)

se.bands <- exp(se.bands.logit) / (1 + exp(se.bands.logit))

```

Alternatively:

```{r}
pred <- predict(fit, newdata = list(age = age.grid), type = "response", se = T)

with(wage, {
   plot(age, I(wage > 250), xlim = agelims, type = "n")
   points(jitter(age), I((wage > 250)/5), cex = .5, pch = "|", col = "darkgrey")
   lines(age.grid, pfit, lwd = 2, col = "blue")
   matlines(age.grid, se.bands, lwd = 1, col = "blue", lty = 3)
})
```

```{r}
table(cut(wage$age, 4))
```

```{r}
fit <- lm(wage ~ cut(age, 4), data = wage)
coef(summary(fit))
```

### Splines

```{r}
fit <- lm(wage ~ bs(age, knots = c(25, 40, 60)), data = wage)

pred <- predict(fit, newdata = list(age = age.grid), se = T)

with(wage, {
   plot(age, wage, col = "gray")
   lines(age.grid, pred$fit, lwd=2)
   lines(age.grid, pred$fit+2*pred$se.fit, lty="dashed")
   lines(age.grid, pred$fit-2*pred$se.fit, lty="dashed")
})
```

```{r}
dim(bs(wage$age, knots = c(25, 40, 60)))
dim(bs(wage$age, df = 6))
attr(bs(wage$age, df = 6), "knots")
```

```{r}
fit2 <- lm(wage ~ ns(age, df = 4), data = wage)
pred2 <- predict(fit2, newdata =  list(age = age.grid), se = T)
par(mfrow=c(1,1))
plot(wage$age, wage$wage, col = "gray")
lines(age.grid, pred2$fit, col = "red", lwd = 2)
```

```{r}
with(wage,{
   plot(age, wage, xlim = agelims, cex = .5, col = "darkgrey")
   title("Smoothing Spline")
   fit <- smooth.spline(age, wage, df = 16)
   fit2 <- smooth.spline(age, wage, cv = T)
   
   lines(fit, col = "red", lwd = 2)
   lines(fit2, col = "blue", lwd = 2)
})
```

```{r}
with(wage, {
   plot(age, wage, xlim = agelims, cex = .5, col = "darkgrey")
   title("Local Regression")
   fit <- loess(wage ~ age, span = .2)
   fit2 <- loess(wage ~ age, span = .5)
   lines(age.grid, predict(fit, data.frame(age = age.grid)), col = "red", lwd = 2)
   lines(age.grid, predict(fit2, data.frame(age = age.grid)), col = "blue", lwd = 2)
   legend("topright", legend = c("Span = 0.2", "Span = 0.5"), col = c("red", "blue"), lty = 1, lwd = 2, cex = .8)
})
```

### GAMs

```{r}
gam1 <- lm(wage ~ ns(year, 4) + ns(age, 5) + education, data = wage)

gam.m3 <- gam(wage ~ s(year, 4) + s(age, 5) + education, data = wage)
par(mfrow = c(1, 3))
plot.Gam(gam.m3, se = T)
```

```{r}
par(mfrow = c(1, 3))
plot(gam.m3, se = T, col = "blue")

plot.Gam(gam1, se = T, col = "red")
```

```{r}
gam.m1 <- gam(wage ~ s(age, 5) + education, data = wage)
gam.m2 <- gam(wage ~ year + s(age, 5) + education, data = wage)

anova(gam.m1, gam.m2, gam.m3)
```

```{r}
summary(gam.m3)
```

```{r}
pred <- predict(gam.m2, newdata = wage)

gam.lo <- gam(wage ~ s(year, df = 4) + lo(age, span = 0.7) + education, data = wage)
plot.Gam(gam.lo, se = T, col = "green")
```

```{r}
plot(gam.lo)
```

```{r}
gam.lr <- gam(I(wage > 250) ~ year + s(age, df = 5) + education, family = binomial, data = wage)
par(mfrow = c(1, 3))
plot(gam.lr, se = T, col = "green")
```

```{r}
table(wage$education, I(wage$wage > 250))
```

```{r}
levels(wage$education)

gam.lr.s <- gam(I(wage > 250) ~ year + s(age, df = 5) + education, family = binomial, data = wage, subset = (education != "1. < HS Grad"))
plot(gam.lr.s, se = T, col = "green")
```

### Applied

In this exercise, you will further analyze the wage data set considered throughout this chapter.

```{r}
test.size <- .7
index <- sample(nrow(wage), nrow(wage) * test.size, replace = F)

train <- wage[index]
test <- wage[!index]

```

a.) Perform polynomial regression to predict wage using age. Use cross-validation to select the optimal degree d for the polynomial. What degree was chosen, and how does this compare to the result of hypothesis testing using ANOVA? Make a plot of the fit obtained.

```{r}
degree <- 20; folds = 10
cv.errors <- numeric(degree)

fold.size <- nrow(train) / folds

for(deg in 1:degree)
{
   # 10 fold cv
   errors <- numeric(folds)
   for(fold in 1:folds)
   {
      holdout <- seq((fold - 1) * fold.size, fold * fold.size)
      
      cv.train <- train[!holdout]
      cv.test <- train[holdout]
      
      fit <- lm(wage ~ poly(age, deg), data = cv.train)
      
      pred <- predict(fit, newdata = cv.test, type = "response")
      
      errors[fold] <- sqrt(mean((cv.test$wage - pred)^2))
   }
   cv.errors[deg] <- mean(errors)
}

lowest.error <- which.min(cv.errors)

cv.results <- data.table(degree = 1:degree, error = cv.errors)[, lowest := degree == lowest.error]

ggplot(cv.results, aes(degree, error, fill = lowest)) +
   geom_bar(stat = "identity") +
   labs(title = "RMSE by Degree")
```

```{r}
model <- lm(wage ~ poly(age, lowest.error), data = train)

test %>%
    mutate(predictions = predict(model, test)) %>%
    ggplot(aes(age, wage, col = 'darkgrey')) +
    geom_point(alpha = .65) +
    geom_line(aes(age, predictions, col = 'cornflowerblue'), size = 1.5) +
    scale_color_manual(name = 'Value Type',
                         labels = c('Observed', 'Predicted'),
                         values = c('cornflowerblue', 'darkgrey' )) +
    labs(x = 'Age', y = 'Wage', 
         title = paste0('Predictions from polynomial model of degree ', lowest.error))
```

b.) Fit a step function to predict wage using age, and perform cross-validation to choose the optimal number of cuts. Make a plot of the fit obtained.

```{r}
cuts <- 20; folds = 10
cv.errors <- numeric(degree)

fold.size <- nrow(train) / folds

for(cuts in 2:cuts)
{
   # 10 fold cv
   errors <- numeric(folds)
   
   # apply cut here so CV train/test have same levels
   train$AgeGroup <- cut(train$age, cuts)
   
   for(fold in 1:folds)
   {
      holdout <- seq((fold - 1) * fold.size, fold * fold.size)
      
      cv.train <- train[!holdout]
      cv.test <- train[holdout]
      
      fit <- lm(wage ~ I(AgeGroup), data = cv.train)
      
      pred <- predict(fit, newdata = cv.test, type = "response")
      
      errors[fold] <- sqrt(mean((cv.test$wage - pred)^2))
   }
   
   cv.errors[cuts] <- mean(errors)
}

lowest.error <- which.min(cv.errors[cv.errors != 0])

cv.results <- data.table(cuts = 1:cuts, error = cv.errors)[, lowest := cuts == lowest.error]

ggplot(cv.results, aes(cuts, error, fill = lowest)) +
   geom_bar(stat = "identity") +
   labs(title = "RMSE by Age Group")
```

```{r}
wage.grouped <- wage
wage.grouped$AgeGroup <- cut(wage.grouped$age, lowest.error)

test.size <- .7
index <- sample(nrow(wage), nrow(wage) * test.size, replace = F)

train <- wage.grouped[index]
test <- wage.grouped[!index]

model <- lm(wage ~ I(AgeGroup), data = train)

test %>%
    mutate(predictions = predict(model, test)) %>%
    ggplot(aes(age, wage, col = 'darkgrey')) +
    geom_point(alpha = .65) +
    geom_line(aes(age, predictions, col = 'cornflowerblue'), size = 1.5) +
    scale_color_manual(name = 'Value Type',
                         labels = c('Observed', 'Predicted'),
                         values = c('cornflowerblue', 'darkgrey' )) +
    labs(x = 'Age', y = 'Wage', 
         title = paste0('Predictions from polynomial model of age group ', lowest.error))
```

The wage data set contains a number of other features not explored in this chapter, such as marital status (*marit1*), job class (*jobclass*), and others. Explore the relationships between some of these other predictors and wage, and use non-linear fitting techniques in order to fit flexible models to the data. Create plots of the results obtained, and write a summary of your findings.

```{r}

```

